An Algebraic Theory File Synchronization

### 2015-12-25 2015/algebra.tex
Further research, conflict comparison, renaming unrelated to incomparable
Improved text
Reordering; improvements; new notation; improved figure

\subsection{Towards an Algebra}\label{sec_algebra}

@@

In this algebra, equivalence ($\equiv$) and extension ($\eqext$) would become relations,
and logical rules involving them
(e.g. $ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $) would be re-cast as inference rules.
In addition to these, the \namecrefs{ax_separate_commute} listed in \cref{section_axioms}
would become axioms that we accept as true, and from which other true statements can be derived
using the inference rules.
Such a system can allow us to deduce, for example, whether two sequences are equivalent
or one extends the other
without reference to filesystems.

@@

In a more complete algebra from which the correctness and completeness
of reconciliation could also be derived,
one can also include symbols, inference rules and axioms for 
$\works{\seqset{X}}$ and $\worksc{\seqset{X}}{\seqset{Y}}$.
We see proving the soundness and completeness of this extended algebra
an intriguing problem that is worthy of further research.

>>>>

\section{Towards a Formal Algebra}\label{sec_algebra}

@@

In this algebra, equivalence ($\equiv$) and extension ($\eqext$) would become relations
between sequences, and logical rules involving them
(e.g. $ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $) would be re-cast as inference rules.
In addition to these, the \namecrefs{ax_separate_commute} listed in \cref{section_axioms}
would become axioms that we accept as true, and from which other true statements can be derived
using the inference rules.
Such a system can allow us to deduce, for example, whether two sequences are equivalent
or one extends the other
without reference to filesystems.

@@

In a more complete algebra from which the correctness and completeness
of reconciliation could also be derived,
one can also include symbols, inference rules and axioms for 
$\works{\seqset{A}}$ and $\worksc{\seqset{A}}{\seqset{B}}$.
We see proving the soundness and completeness of this extended algebra
an intriguing problem that is worthy of further research.

### 2015-12-25 2015/axioms.tex
Reordering; improvements; new notation; improved figure

% TODO Ref back to intro; further abstraction - no need to go the FS. break-all, equiv, extend

So that we could describe the effects of commands on a filesystem, let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we will investigate sequences of commands with a well-defined order.
\begin{mydef}[$\cc$ and $\emptyseq$]
We use ``$\cc$'' to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbdaa{n}{\vald}\cc\cbfaa{m}{\valf}$, or $S'=S\cc\cbba{o}$,
with the meaning that the command on the left is applied first:
\[ \alpha\cc\beta\FS = \beta(\alpha\FS). \]
Sequences of commands form a free semigroup, and, as usual,
$\emptyseq$ denotes its unit element, the empty sequence,
which by definition leaves all filesystems unchanged.
\end{mydef}

@@

\begin{mydef}[$\equiv$]
% Relation in the algebra
$A\equiv B$ means that $A$ and $B$ are equivalent,
% intended interpretation:
that is, they behave in the same way on
all filesystems: $\forall \FS: A\FS=B\FS$.
It is a reflexive, commutative and transitive relation,
and clearly
% inference rule:
$ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $.
\end{mydef}

\begin{mydef}[$\eqext$]
% Relation in the algebra
$A\eqext B$ means that $B$ extends $A$,
% intended interpretation:
that is, they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS: A\FS\neq\fsbroken\Rightarrow A\FS=B\FS$.
It is a reflexive and transitive relation,
and we can see that
% inference rules:
$ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$
and
$ A\equiv B \Rightarrow A\eqext B$.
\end{mydef}

@@

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
$\cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n}$ 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
(Note that this and next rule is true because the assertion command is
adjacent to the original command.)
\end{myax}

\begin{myax}\label{ax_parent_assert}
An assertion command can be added on an ancestor node:
$\cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn}$ 
where $\nn\descendant n$ and $\cxy\neq\cbb$.
\end{myax}

@@

>>>>

So that we could describe the effects of commands in general,
let us introduce some notation
and note some observations.
We already know that
commands usually do not occur in isolation,
and are applied to filesystems in time.
Therefore we will investigate sequences of commands with a well-defined order.
\begin{mydef}[$\cc$ and $\emptyseq$]
We use ``$\cc$'' to concatenate commands to form a sequence, or concatenate sequences to form a longer sequence:
$S=\cbdaa{n}{\vald}\cc\cbfaa{m}{\valf}$, or $S'=S\cc\cbba{o}$,
with the meaning that the command on the left is applied first:
\[ \alpha\cc\beta\aFS = \beta(\alpha\aFS). \]
Sequences of commands form a free semigroup, and, as usual,
$\emptyseq$ denotes its unit element, the empty sequence,
which by definition leaves all filesystems unchanged.
\end{mydef}

@@

\begin{mydef}[$\equiv$]
% Relation in the algebra
$A\equiv B$ means that $A$ and $B$ are equivalent,
% intended interpretation:
that is, they behave in the same way on
all filesystems: $\forall \FS: A\aFS=B\aFS$.
It is a reflexive, commutative and transitive relation,
and clearly
% inference rule:
$ A\equiv B \Rightarrow S\cc A\cc T\equiv S\cc B\cc T $.
\end{mydef}

\begin{mydef}[$\eqext$]
% Relation in the algebra
$A\eqext B$ means that $B$ extends $A$,
% intended interpretation:
that is, they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS: A\FS\neq\fsbroken\Rightarrow A\aFS=B\aFS$.
It is a reflexive and transitive relation,
and we can see that
% inference rules:
$ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$
and
$ A\equiv B \Rightarrow A\eqext B$.
\end{mydef}

@@

\begin{myax}\label{ax_child_assert}
An assertion command can be added on a descendant node:
\[ \cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n} \] 
where $\nn\descendant n$ and $\cxy\neq\cdd$.
(Note that this and next rule is true because the assertion command is
adjacent to the original command.)
\end{myax}

\begin{myax}\label{ax_parent_assert}
An assertion command can be added on an ancestor node:
\[ \cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn} \]
where $\nn\descendant n$ and $\cxy\neq\cbb$.
\end{myax}

@@

% TODO We will only use these except for...

\noindent
We also define the concept of two commands being independent:

\begin{mydef}[$A\indep B$: Independent commands, sequences and sets]
Two commands $\cxynv$ and $\czwmv$ 
are independent, and we write $\cxynv\indep\czwmv$ if 
they have the same effect in either order and do not break all filesystems:
\[ \cxynv\cc\czwmv \equiv \czwnv\cc\cxynv \nequiv \cbrk. \]
For two sequences or unordered sets of commands $A$ and $B$ we write $A\indep B$ if
for all $\alpha$ in $A$ and all $\beta$ in $B$, $\alpha\indep\beta$.
\end{mydef}

It is intentional that we use the same symbol for independent commands
and sequences as for incomparable nodes. As we will in the next
\namecref{incomparable_is_independent},
these two concepts are closely related.

\begin{mylem}\label{incomparable_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are incomparable:
\[ \cxynv\indep\czwmv \Leftrightarrow n\unrel m \]
if $\cxynv\neq\czwmv$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
This is easy to show based on the output of the software that investigates the behavior of command pairs.
However, this proposition can also be derived from the \namecrefs{ax_separate_commute} already listed
in the following way.

We proceed in two steps. First, we note that
from \cref{ax_separate_commute,ax_separate_nobreaks} we know that
if $n\unrel m$, then $\cxynv$ and $\czwmv$ must necessarily be independent, and therefore
$ \cxynv\nindep\czwmv \wedge n\unrel m $ cannot be the case.

Next, we prove that
\[ \cxynv\indep\czwmv \wedge n\nunrel m \] is also impossible as it would lead to contradiction.
As $n\nunrel m$
we know that $n=m$, $n\descendant m$ or $m\descendant n$.

If $n=m$, then from \cref{ax_same_breaks} we know $Y=Z$ and $W=X$
as otherwise the commands, in one order or the other, would break all filesystems
and they could not be independent.
The commands are not assertion commands, and so from \cref{ax_same_singlec}
we know that $\cxynv\cc\czwnv\equiv\czwnv\cc\cxynv$ can only happen if $Y=W$ and $\valvy=\valvw$.
Therefore $X=Y=Z=W$ which is only possible if both commands are $\cff$ and have the same
output value. This, however, contradicts our assumption that they are different.

If $n\descendant m$ or $m\descendant n$, then
from \cref{ax_distantrel_breaks} we know that if they are not directly related,
then $\cxynv\cc\czwmv$ breaks all filesystems, and they cannot be independent.
From the construction and destruction pairs and 
\cref{ax_directchild_breaks,ax_directparent_breaks} we also see that
even if they are directly related, either
$\cxynv\cc\czwmv$ or $\czwmv\cc\cxynv$ 
breaks all filesystems, so they again cannot be independent.
\end{proof}

### 2015-12-25 2015/conclusion.tex
Reordering; improvements; new notation; improved figure

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{centernot}
\usepackage{tikz}
\usepackage{cleveref}

\input{macros}

\title{Complete Algebraic File Synchronization}
\author{Elod Pal Csirmaz}

\begin{document}
\maketitle
\begin{abstract}
Abstract goes here
\end{abstract}

\input{intro}

\section{Definitions}\label{sec_def}

\input{def-filesystem}

\input{def-commands}

\input{axioms}

\section{Update Detection}\label{sec_update}

\input{upd-def}

\input{ordering}

\input{upd-correct}

\section{Reconciliation}\label{sec_rec}

\input{rec-intro}

\input{works-inverse}

\input{rec-sound}

\input{rec-complete}

\input{comparisons}

\input{algebra}

@@

% TODO
Conclusions go here --- same approach to other systems
\medskip

@@

Apart from constructing an algebra,
there are many other ways in which further work can extend the current results.
We think one of the most interesting extensions would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and completeness of the algorithm proposed.
A related problem is to model cases where reconciliation cannot
complete fully (due to a system or network failure, for example)
or if only a subset of the replicas are reconciled,
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystem.
Identifying the information one needs to keep to perform
update detection in such cases and devising a suitable algorithm
would be beneficial for any implementation of a reconciler
based on the ideas presented here.

@@

\section{Acknowledgments}

I thank Laszlo Csirmaz for his invaluable input on previous drafts of this paper,
and Bill Zissimopoulos for his observations on algebraic file synchronization that triggered this research.

\begin{thebibliography}{99}

\input{references}

\end{thebibliography}

\end{document}

>>>>

@@

% TODO
... ... ...

(Same approach to other systems)
\medskip

@@

Apart from constructing an algebra,
there are many other ways in which further work can extend the current results.
We think one of the most interesting extensions would be to
consider reconciling not only two, but more replicas in a single step and
prove the correctness and maximality of the algorithm proposed.
A related problem is to model cases where reconciliation cannot
complete fully (due to a system or network failure, for example)
or if only a subset of the replicas are reconciled,
both of which would result in a state where different replicas
have different common ancestors, that is,
the updates specific to the replicas start from different points
in the update history of the filesystem.
Identifying the information one needs to keep to perform
update detection in such cases and devising a suitable algorithm
would be beneficial for any implementation of a reconciler
based on the ideas presented here.

@@

### 2015-12-25 2015/def-commands.tex
Reordering; improvements; new notation; improved figure

@@

Let us consider what kind of information is usually encoded in filesystem operations.
A minimal set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$, which creates a file or directory ($\valv$) at the empty node $n$;
$edit(n,\valv)$, which replaces the earlier value at $n$ with $\valv$;
and $remove(n)$, which deletes the value at $n$---where 
$n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work 
demonstrated\footnote{Bill Zissimopoulos, personal communication, August 2015.}
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

@@

We therefore have four pieces of information related to each command:
its input type, its output type, the node to which it is applied, and finally
the value that it stores at the node.
We will denote commands using quadruples that list these
in a the following format:
\[ \cxynv, \]
where $X$ is the input type, $Y$ is the output type, $n$ is the node,
and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
We use $\cchard$, $\ccharf$ or $\ccharb$ to represent the different
types. For example, $\cbdaa{n}{\vald}$ 
will represent $mkdir(n,\vald)$.
Alternatively, we use a triple of the two types and a node if 
there is only a single suitable output value for the given type (e.g. $\empt$).
That is, $\cdba{n}$ will represent $rmdir(n)$.

@@

We note that a command or a sequence of commands is applied to a filesystem
by prefixing the command or sequence to it, for example: $\cbrk\FS$, $\cbdaa{n}{\vald}\FS$, 
or simply $S\FS$ if $S$ is a sequence of commands. % TODO Add space between command and FS

@@

\begin{mydef}[Replacement operator]
Using $\FS[V/n]$ we denote a filesystem derived from $\FS$:
\[ \FS[V/n](m) :=
   \begin{cases}
   V &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
\end{mydef}

The effect of the commands is as follows:
\begin{itemize}
\item $\cbrk\FS = \fsbroken$
\item
\[ \cxynv\FS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[V/n] \mbox{~violates the tree property}\\
   \FS[V/n] &\mbox{otherwise.}
   \end{cases}
\]
\end{itemize}

>>>>

% TODO identity command?

@@

Let us consider what kind of information is usually encoded in filesystem operations.
A minimal set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$, which creates a file or directory ($\valv$) at the empty node $n$;
$edit(n,\valv)$, which replaces the earlier value at $n$ with $\valv$;
and $remove(n)$, which deletes the value at $n$---where 
$n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work
demonstrated (\cite{BZ})
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

@@

We therefore have four pieces of information related to each command:
its input type, its output type, the node to which it is applied, and finally
the value that it stores at the node.
We will denote commands using quadruples that list these
in a the following format:
% TODO convert into definition
\[ \cxynv, \]
where $X$ is the input type, $Y$ is the output type, $n$ is the node,
and $\valvy$ is a value of the appropriate type from $\setvx{Y}$.
We use $\cchard$, $\ccharf$ or $\ccharb$ to represent the different
types. For example, $\cbdaa{n}{\vald}$ 
will represent $mkdir(n,\vald)$.
% TODO Just say "omit"
Alternatively, we use a triple of the two types and a node if 
there is only a single suitable output value for the given type (e.g. $\empt$).
That is, $\cdba{n}$ will represent $rmdir(n)$.

@@

We note that a command or a sequence of commands is applied to a filesystem
by prefixing the command or sequence to it, for example: $\cbrk\aFS$, $\cbdaa{n}{\vald}\aFS$, 
or simply $S\aFS$ if $S$ is a sequence of commands.

@@

% TODO include command in def
\begin{mydef}[Replacement operator]
Using $\FS[\valv/n]$ we denote a filesystem derived from $\FS$:
\[ \FS[\valv/n](m) :=
   \begin{cases}
   \valv &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
\end{mydef}

The effect of the commands is as follows:
\begin{align*}
&\cbrk\aFS = \fsbroken \\
&\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} \FS(n)\not\in\setvx{X}\\
   \fsbroken &\mbox{if~} \FS[\valvy/n] \mbox{~violates the tree property}\\
   \FS[\valvy/n] &\mbox{otherwise.}
   \end{cases}
\end{align*}

### 2015-12-25 2015/intro.tex
Reordering; improvements; new notation; improved figure

In this paper we investigate an operation- or command-based approach
to filesystem synchronization.
Presented with multiple copies (replicas) of a filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce (\cite{BP}):
update detection, where we identify modifications that have been applied to the replicas
since they diverged,
and reconciliation, where we identify updates that can be propagated to other replicas and do so.
The remaining updates, which could not be propagated, are conflicting updates;
resolving these requires a separate system or human intervention.

@@

\Cref{sec_rec} defines our reconciliation 
and conflict detection algorithms.
We then proceed to prove that the output of reconciliation
is sound in the sense that 
it can be applied to the replicas without causing an error,
and it is complete
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of \emph{conditional operation} of sequences of commands,
and show that it has a number of highly convenient properties.
We finish by comparing our results to other research
and outlining directions for further work
on the introduced algebraic system.

The approach in the present paper offers an improvement
over results presented in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates.
This has the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as 
predictions made by the reasoning are more accurate
due to the additional information content.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it complete.

>>>>

In this paper we investigate an operation- or command-based approach
to filesystem synchronization.
Presented with multiple copies (replicas) of the same filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce (\cite{BP}):
update detection, where we identify modifications that have been applied to the replicas
since they diverged,
and reconciliation, where we identify updates that can be propagated to other replicas and do so.
The remaining updates, which could not be propagated, are conflicting updates;
resolving these requires a separate system or human intervention.

@@

\Cref{sec_rec} defines our reconciliation 
and conflict detection algorithms.
We then proceed to prove that the output of reconciliation
is correct in the sense that 
it can be applied to the replicas without causing an error,
and it is maximal
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of \emph{conditional operation} of sequences of commands,
and show that it has a number of highly convenient properties.
We finish by comparing our results to other research
and outlining directions for further work
on the introduced algebraic system.

The approach in the present paper offers an improvement
over results presented in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates.
This has the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as 
predictions made by the reasoning are more accurate
due to the additional information content.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it maximal.

### 2015-12-25 2015/rec-complete.tex
More text improvements
Reordering; improvements; new notation; improved figure

\subsection{The Completeness of Reconciliation}

@@

The reconciliation algorithm defined above is also complete, that is,
it is not possible to apply any further commands from $A\setminus B$ to $\FS_B$.
To show this, we are going to prove the following \namecref{rec_is_complete}:

\begin{myth}\label{rec_is_complete}
If $A$ and $B$ are simple sequences,
and we select a sequence $S\subset \amb$ so that
it contains a command $\cxynv$ for which $\cxynv\nindep\bma$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ do not break
(that is, $S$ breaks all possible $\FS_B$ replicas),
or it changes a node that $B$ has already changed to a different value,
that is, it overrides a change in $\FS_B$.
\end{myth}
Such an override could occur if a given node was modified differently in
$A$ and $B$ (to different values but the same type), which is clearly
a conflict that should be marked for review
by a human or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $\bma$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep\bma$.
Let us also split $\bma$ into $T_0\cc\czwmv\cc T_1$ where $T_1\indep\cxynv$.
From \cref{can_move_intersection} we know that
$B\cc S \equiv \acb\cc\bma\cc S$.
We also know that commands in $S_0$ and $\bma$ commute, and so this is equivalent to
$\acb\cc S_0\cc\bma\cc\cxynv\cc S_1$.
Expanding $\bma$ and swapping $T_1$ and $\cxynv$ we get
\[ \acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv\cc T_1\cc S_1. \]

First, we prove that this sequence breaks all filesystems
unless $n=m$.
Let us therefore suppose $n\neq m$ and, to use an inverse proof,
that the sequence works on $\FS$ where $A\FS\neq\fsbroken$.
If so, its prefix, $\acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv$ must also work.

We know $\cxynv\nindep\czwmv$, and so from \cref{incomparable_is_independent}, $n\nunrel m$.
As the sequence works, $\czwmv\cc\cxynv$ can only be a construction or destruction pair.
As $\czwmv$ and $T_0$ originate from $B\setminus A$ and $B$ is simple, we also know there are
no commands on $m$ in $A\cap B$ or $T_0$,
and that there are no commands on $m$ in $S_0$ as $S_0\indep\bma$.

@@

Clearly $Q=Z$ (from \cref{worksinputmatch}), and from the construction and destruction
pairs we also see that $R=W$ must hold, where $W=\cchard$ or $W=\ccharb$.
As $|\setd|=1$ and $|\setb|=1$, this also means that $v_R=v_W$ and therefore
$\cqrmv = \czwmv$.
This is a contradiction as $\cqrmv\in A$, but $\czwmv\in\bma$.

The only possibility is therefore $n=m$.
As $A$ and $B$ are simple, we therefore know that these are the first commands on $n$,
and so (again from \cref{worksinputmatch}) $Z=X$.
We also know $\valvw\neq\valvy$ as otherwise the two commands would be equal and
would be included in $A\cap B$,
but this means that $\cxynv$, from $A$, overrides a change introduced 
by $\czwnv$, from $B$, which should be treated as a conflict.
\end{proof}

>>>>

\subsection{Reconciliation is Maximal}

@@

The reconciliation algorithm defined above is also maximal, that is,
it is not possible to apply any further commands from $A\setminus B$ to $\FS_B$.
To show this, we are going to prove the following \namecref{rec_is_complete}:

\begin{myth}\label{rec_is_complete}
If $A$ and $B$ are simple sequences,
and we select a sequence $S\subset \amb$ so that
it contains a command $\cxynv$ for which $\cxynv\nindep\bma$,
then $B\cc S$ either breaks all filesystems $A$ and $B$ do not break
(that is, $S$ breaks all possible $\FS_B$ replicas),
or it changes a node that $B$ has already changed to a different value,
that is, it overrides a change in $\FS_B$.
\end{myth}
Such an override could occur if a given node was modified differently in
$A$ and $B$ (to different values but the same type), which is clearly
a conflict that should be marked for review
by a human or a different system.
\begin{proof}
The proof is similar to the ones we have seen above.
Without loss of generality, we assume that $\cxynv$ is the first command in $S$
that is not independent of $\bmanp$.
If so, we can split $S$ into $S_0\cc\cxynv\cc S_1$ where $S_0\indep\bma$.
Let us also split $\bmanp$ into $T_0\cc\czwmv\cc T_1$ where $T_1\indep\cxynv$.
From \cref{can_move_intersection} we know that
$B\cc S \equiv \acb\cc\bma\cc S$.
We also know that commands in $S_0$ and $\bmanp$ commute, and so this is equivalent to
$\acb\cc S_0\cc\bma\cc\cxynv\cc S_1$.
Expanding $\bmanp$ and swapping $T_1$ and $\cxynv$ we get
\[ \acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv\cc T_1\cc S_1. \]

First, we prove that this sequence breaks all filesystems
unless $n=m$.
Let us therefore suppose $n\neq m$ and, to use an inverse proof,
that the sequence works on $\FS$ where $A\FS\neq\fsbroken$.
If so, its initial segment, $\acb\cc S_0\cc T_0 \cc \czwmv \cc \cxynv$ must also work.

We know $\cxynv\nindep\czwmv$, and so from \cref{incomparable_is_independent}, $n\nunrel m$.
As the sequence works, $\czwmv\cc\cxynv$ can only be a construction or destruction pair.
As $\czwmv$ and $T_0$ originate from $\bmanp$ and $B$ is simple, we also know there are
no commands on $m$ in $\acbnp$ or $T_0$,
and that there are no commands on $m$ in $S_0$ as $S_0\indep\bma$.

@@

Clearly $Q=Z$ (from \cref{worksinputmatch}), and from the construction and destruction
pairs we also see that $R=W$ must hold, where $W=\cchard$ or $W=\ccharb$.
As $|\setd|=1$ and $|\setb|=1$, this also means that $v_R=v_W$ and therefore
$\cqrmv = \czwmv$.
This is a contradiction as $\cqrmv\in A$, but $\czwmv\in\bmanp$.

The only possibility is therefore $n=m$.
As $A$ and $B$ are simple, we therefore know that these are the first commands on $n$,
and so (again from \cref{worksinputmatch}) $Z=X$.
We also know $\valvw\neq\valvy$ as otherwise the two commands would be equal and
would be included in $\acbnp$,
but this means that $\cxynv$, from $A$, overrides a change introduced 
by $\czwnv$, from $B$, which should be treated as a conflict.
\end{proof}

### 2015-12-25 2015/rec-intro.tex
Reordering; improvements; new notation; improved figure

If we start with two copies of the filesystem $\FS$,
and two different sequences are applied to the copies to yield $\FS_A:=A\FS$
and $\FS_B:=B\FS$, then our aim is to define sequences of commands $\reca$ and $\recb$
so that $\recb\FS_A$ and $\reca\FS_B$ would be as close to each other as possible.

We work based on the assumption that to achieve this, we need
to apply to $\FS_B$ the commands that have been applied to $\FS_A$, and \emph{vice versa}.
As some commands may have been applied to both filesystems, our first approximation
is $\reca = A\setminus B$ and $\recb = B\setminus A$
(in some suitable order).
This, however, will break both filesystems if there have been incompatible updates
in $A$ and $B$. 
Our aim is therefore to provide an algorithm that selects the commands 
$\reca \subset A\setminus B$
and $\recb \subset B\setminus A$ 
so that $\reca\FS_B\neq\fsbroken$ and $\recb\FS_A\neq\fsbroken$,
and show that these are the longest sequences with this property, that is,
adding any command left out from $\reca$ or $\recb$ will break the filesystems.

We assume that $A$ and $B$ are returned by an update detector,
and so they are simple sequences.
(Otherwise, they can be simplified to become simple sequences; see \cref{update_works}.)
We also assume that $\FS_A$ and $\FS_B$ are not broken.
With these assumptions,
we can now define independent sequences and
our reconciliation algorithm.

\begin{mydef}{Independent commands, sequences and sets.}
Two commands $\cxynv$ and $\czwmv$ 
are independent, and we write $\cxynv\indep\czwmv$ if 
they have the same effect in either order and do not break all filesystems:
$\cxynv\cc\czwmv \equiv \czwnv\cc\cxynv \nequiv \cbrk$.
For two sequences or sets of commands $A$ and $B$ we write $A\indep B$ if
$\forall \alpha\in A \mbox{~and~} \forall \beta\in B: \alpha\indep\beta$.
\end{mydef}

\begin{mydef}{Reconciliation.}\label{def:reconciliation}
\begin{align*}
\reca = \{\alpha| & \\
& \alpha\in A\setminus B \\
& \wedge \\
& \forall \beta\in B\setminus A: \alpha\indep\beta\\
\} &
\end{align*}

In other words, the sequence $\reca$ we apply to $\FS_B$ 
to reconcile it with $\FS_A$
is the largest subset of $A\setminus B$
that is independent of $B\setminus A$.
The order of commands in the sequence can be determined using the algorithm described in
\cref{ordering} as any subset of $A$ is also simple.
In a similar way, $\recb$ can be obtained by reversing $A$ and $B$
in the definition.
\end{mydef}

We would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\FS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
let us introduce two relations that describe under what conditions
sequences of commands work, that is, do not break a filesystem.

>>>>

If we start with two copies of the filesystem $\FS$,
and two different sequences are applied to the copies to yield $\FS_A:=A\aFS$
and $\FS_B:=B\aFS$, then our aim is to define sequences of commands $\reca$ and $\recb$
so that $\recb\aFS_A$ and $\reca\aFS_B$ would be as close to each other as possible.

We work based on the assumption that to achieve this, we need
to apply to $\FS_B$ the commands that have been applied to $\FS_A$, and \emph{vice versa}.
As some commands may have been applied to both filesystems, our first approximation
is $\reca = A\setminus B$ and $\recb = B\setminus A$
(in some suitable order).
This, however, will break both filesystems if there have been incompatible updates
in $A$ and $B$. 
Our aim is therefore to provide an algorithm that selects the commands 
$\reca \subset A\setminus B$
and $\recb \subset B\setminus A$ 
so that $\reca\aFS_B\neq\fsbroken$ and $\recb\aFS_A\neq\fsbroken$,
and show that these are the longest sequences with this property, that is,
adding any command left out from $\reca$ or $\recb$ will break the filesystems.

We assume that $A$ and $B$ are returned by an update detector,
and so they are simple sequences.
(Otherwise, they can be simplified to become simple sequences; see \cref{update_works}.)
We also assume that $\FS_A$ and $\FS_B$ are not broken.
With these assumptions,
we can now define our reconciliation algorithm.

\begin{mydef}[Reconciliation]\label{def:reconciliation}
We determine $\reca$---which we apply to $\FS_B$ 
to reconcile it with $\FS_A$---to be
the largest subset of $A\setminus B$
that is independent of $B\setminus A$:
\begin{align*}
\reca = \{\alpha|{}& \\
& \alpha\in A\setminus B \\
& \wedge \\
& \forall \beta\in B\setminus A: \alpha\indep\beta\\
\}. &
\end{align*}
The order of commands in the sequence can be determined using the algorithm described in
\cref{ordering} as any subset of $A$ is also simple.
In a similar way, $\recb$ can be obtained by reversing $A$ and $B$
in the definition.
\end{mydef}

We would now like to prove that $\reca$ and $\recb$ can be applied to the replicas,
that is, without loss of generality,
\[ \reca\aFS_B\neq\fsbroken. \]
So that we can formalize statements needed to prove this,
let us introduce two relations that describe under what conditions
sequences of commands work, that is, do not break a filesystem.
